---
title: "Assignment"
author: "Priscilla Hare and Callie Steffen"
date: "5/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load the libraries into this R session

library(raster)       #Main raster library with nearly all functions used in this analysis
library(rgdal)        #Spatial library - most functions used from rgdal are for vectors (shapefiles)
library(rasterVis)    #Useful for raster visualizations
library(maps)         #Has a database of maps. I use this to add a map to my raster to visualize land boundaries
library(rgeos)        #Need this library for topology operations on geometries
library(dplyr)        #NOT spatial - this is a data wrangling library
library(RColorBrewer) #Also not spatial - used to set the spectral color scheme 
library(tmap)

```

```{r}
# rainbow color scheme
cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) 

#setting smaller margins for plotting
par(mar=c(2,2,1,1))
```


Load threat layers for climate: acid, sst, uv. and load species
```{r}

acid <- raster("Threats_data/impact_acid.tif")
#plot(acid, col = cols)

sst <- raster("Threats_data/impact_sst.tif")
#plot(sst, col = cols)

uv <- raster("Threats_data/impact_uv.tif")
#plot(uv, col = cols)

all_spp <- raster("ca_curr_sp_rich.tif")


```
stack climate layers
```{r}
climate_stack <- stack(acid, sst, uv)
#plot(climate_stack, col = cols)
#resolution and extents already the same
```
reclassify zeroes as NAs if needed
```{r}
#hist(acid)
acid <- reclassify(acid, rcl = c(-Inf, 0, NA))
#plot(acid)

#hist(sst)
#sst_test <- reclassify(sst, rcl = c(-Inf, 0 , NA))
sst_masked <- mask(sst, acid) #mask sst layer to shape of reclassified acid layer (removes NAs coded as 0 and keep real 0s)
#plot(sst_masked)

#hist(uv)
uv <- reclassify(uv, rcl = c(-Inf, 0, NA))
#plot(uv)

```

Sum climate threats 
```{r}
climate_threats <- overlay(acid, uv, sst_masked, fun = function(x,y,z){x + y + z})

#plot(climate_threats)
```

Resample to get extent and resolution same between threats and species
```{r}

climate_crop <- crop(climate_threats, all_spp) # correct extent for threats

spp_res <- resample(all_spp, climate_crop, method = 'ngb')

#get zeroes out of species
spp_res <- reclassify(spp_res, rcl = c(-Inf,0,NA))

```

Take top 20%
```{r}
#first threats

climate_cutoff <- quantile(climate_crop, 0.8)
climate_max <- cellStats(climate_crop, max)

rcl_mat_cl <- c(-Inf, climate_cutoff, 0,
             climate_cutoff, climate_max, 1)

climate_binary <- reclassify(climate_crop, rcl = rcl_mat_cl)

#plot(climate_binary)

#now species
spp_cutoff <- quantile(spp_res, 0.8)
spp_max <- cellStats(spp_res, max)

rcl_mat_sp <- c(-Inf, spp_cutoff, 0,
                spp_cutoff, spp_max, 1)

spp_binary <- reclassify(spp_res, rcl = rcl_mat_sp)

#plot(spp_binary)

```
Now overlay climate threats and spp top 20
```{r}
climate_hotpots <- overlay(spp_binary, climate_binary, fun = function(x,y){x + y})

climate_hotpots2 <- reclassify(climate_hotpots, rcl = c(-Inf, 1.5, NA))

hotspot_cols <- c("firebrick")

plot(climate_hotpots2, col = hotspot_cols, legend = F)
map("world", fill = T, add = T, col = 'gray80') #this stopped working. not sure why
```


```{r}

data("World")

hotspot_cols2 <- c("midnightblue")

climate_hotspot_map <- tm_shape(climate_hotpots2) +
  tm_raster("layer", title = " ", palette = hotspot_cols2, labels = "Climate Threat Hotspots") +
  tm_shape(World) +
  tm_fill("gray80") +
  tm_borders("black") +
  tm_layout(bg.color = "lightblue",
            legend.position = c(0.05, .15),
            legend.text.color = "black",
            legend.text.size = .5, 
            legend.width = 5) +
  tm_compass(north = 0, type = "arrow", position = c("right", "top")) +
  tm_credits("Map: Priscilla Hare, June 1, 2019\nSource: Halpern et al 2009\nPublisher: ESM 270", position = c(0.05, .05), size = .4)

climate_hotspot_map

tmap_save(climate_hotspot_map, "climate_hs_map.png", height = 5)
```



## Now do the same for fishing threats


Load threat layers for fishing: recreational, bycatch (high and low), pelagiv (high low), demersal destructive
```{r}

rec <- raster("Threats_data/impact_rec_fish.tif")

dem_nd_hb <- raster("Threats_data/impact_dem_nd_hb.tif")

dem_nd_lb <- raster("Threats_data/impact_dem_nd_lb.tif")

dem_pel_hb <- raster("Threats_data/impact_pel_hb.tif")

dem_pel_lb <- raster("Threats_data/impact_pel_lb.tif")

dem_dem_d <- raster("Threats_data/impact_dem_d.tif")


```

stack climate layers
```{r}
fish_stack <- stack(rec, dem_nd_hb, dem_nd_lb, dem_pel_hb, dem_pel_lb, dem_dem_d)
#plot(fish_stack, col = cols)
#resolution and extents already the same
```
reclassify zeroes as NAs if needed
```{r}

rec_masked <- mask(rec, sst_masked) #mask rec layer to shape of masked sst layer (removes NAs coded as 0 and keep real 0s). because we already like the extent of sst amsked
#plot(rec_masked, col = cols); map('world', fill = T, add = T, col = 'gray80')

dem_nd_hb_masked <- mask(dem_nd_hb, sst_masked)
#plot(dem_nd_hb_masked, col = cols); map('world', fill = T, add = T, col = 'gray80')

dem_nd_lb_masked <- mask(dem_nd_lb, sst_masked)
#plot(dem_nd_lb_masked, col = cols); map('world', fill = T, add = T, col = 'gray80')

dem_pel_hb_masked <- mask(dem_pel_hb, sst_masked)
#plot(dem_pel_hb_masked, col = cols); map('world', fill = T, add = T, col = 'gray80')

dem_pel_lb_masked <- mask(dem_pel_lb, sst_masked)
#plot(dem_pel_lb_masked, col = cols); map('world', fill = T, add = T, col = 'gray80')

dem_dem_d_masked <- mask(dem_dem_d, sst_masked)
#plot(dem_dem_d_masked, col = cols); map('world', fill = T, add = T, col = 'gray80')

```

Sum fishing threats 
```{r}
fishing_threats <- overlay(rec_masked, dem_nd_hb_masked, dem_nd_lb_masked, dem_pel_hb_masked, dem_pel_lb_masked, dem_dem_d_masked, fun = function(a,b,c,d,e,f){a + b + c + d + e + f})

#plot(fishing_threats)
```

Resample to get extent and resolution same between threats and species
```{r}

fishing_crop <- crop(fishing_threats, all_spp) # correct extent for threats

#plot(fishing_crop)

spp_res <- resample(all_spp, fishing_crop, method = 'ngb')

#get zeroes out of species
spp_res <- reclassify(spp_res, rcl = c(-Inf,0,NA))

```

Take top 20%
```{r}
#first threats
fishing_cutoff <- quantile(fishing_crop, 0.8)
fishing_max <- cellStats(fishing_crop, max)

rcl_mat_fish <- c(-Inf, fishing_cutoff, 0,
             fishing_cutoff, fishing_max, 1)

fishing_binary <- reclassify(fishing_crop, rcl = rcl_mat_fish)

#plot(fishing_binary)

#now species
spp_cutoff <- quantile(spp_res, 0.8)
spp_max <- cellStats(spp_res, max)

rcl_mat_sp <- c(-Inf, spp_cutoff, 0,
                spp_cutoff, spp_max, 1)

spp_binary <- reclassify(spp_res, rcl = rcl_mat_sp)

#plot(spp_binary)

```
Now overlay climate threats and spp top 20
```{r}
fishing_hotpots <- overlay(spp_binary, fishing_binary, fun = function(x,y){x + y})

hotspot_cols <- c("lightblue", "lightblue", "firebrick")

#plot(fishing_hotpots, col = hotspot_cols, legend = F)
#map('world', fill = T, add = T, col = 'gray80')

```

```{r}
fishing_hotpots2 <- reclassify(fishing_hotpots, rcl = c(-Inf, 1.5, NA))

hotspot_cols <- c("firebrick")

#plot(climate_hotpots2, col = hotspot_cols, legend = F); map("world", fill = T, add = T, col = 'gray80') #this stopped working. not sure why
```


```{r}

data("World")

fishing_hotspot_map <- tm_shape(fishing_hotpots2) +
  tm_raster("layer", title = " ", palette = hotspot_cols, labels = "Fishing Threat Hotspots") +
  tm_shape(World) +
  tm_fill("gray80") +
  tm_borders("black") +
  tm_layout(bg.color = "lightblue",
            legend.position = c(0.05, .15),
            legend.text.color = "black",
            legend.text.size = .5, 
            legend.width = 5) +
  tm_compass(north = 0, type = "arrow", position = c("right", "top")) +
  tm_credits("Map: Callie Steffen, June 1, 2019\nSource: Halpern et al 2009\nPublisher: ESM 270", position = c(0.05, .05), size = .4)

fishing_hotspot_map

tmap_save(fishing_hotspot_map, "fishing_hs_map.png", height = 5)
```
```{r}

hotspot_cols2 <- c("midnightblue")

overaly_hotspot_map <- tm_shape(fishing_hotpots2) +
  tm_raster("layer", title = " ", palette = hotspot_cols, labels = "Fishing Threat Hotspots") +
  tm_shape(climate_hotpots2) +
  tm_raster("layer", title = " ", palette = hotspot_cols2, labels = "Climate Threat Hotspots", alpha = 0.8) +
  tm_shape(World) +
  tm_fill("gray80") +
  tm_borders("black") +
  tm_layout(bg.color = "lightblue",
            legend.position = c(0.05, .15),
            legend.text.color = "black",
            legend.text.size = .5, 
            legend.width = 5) +
  tm_compass(north = 0, type = "arrow", position = c("right", "top")) +
  tm_credits("Map: Priscilla Hare, June 1, 2019\nSource: Halpern et al 2009\nPublisher: ESM 270", position = c(0.05, .05), size = .4)

overaly_hotspot_map


tmap_save(overaly_hotspot_map, "fishingclimate_hs_map.png", height = 5)

```

