---
title: "CP_Spatial_R"
author: "Priscilla Hare"
date: "5/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Part 1: Setup
```{r}
# Load the libraries into this R session
library(raster)       #Main raster library with nearly all functions used in this analysis
library(rgdal)        #Spatial library - most functions used from rgdal are for vectors (shapefiles)
library(rasterVis)    #Useful for raster visualizations
library(maps)         #Has a database of maps. I use this to add a map to my raster to visualize land boundaries
library(rgeos)        #Need this library for topology operations on geometries
library(dplyr)        #NOT spatial - this is a data wrangling library
library(RColorBrewer) #Also not spatial - used to set the spectral color scheme 
library(tmap)
```

Set Color Palette for Later
```{r}
# view some color palettes
# display.brewer.all()
# rainbow color scheme
cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) 

#setting smaller margins for plotting
par(mar=c(2,2,1,1))
```


### Part 2: Import Raster Data and Visualize
```{r}

all_threats <- raster("Threats_data/full_modelnv.tif")

plot(all_threats, col = cols)

#all_threats
#do the above in console to get info about the raster: dimensions, resolution, extent, coord.ref, data source

# add a landmap to your shapefile. the add=T argument tells R to add it to the existing plot.
# make sure you understand what the other arguments do
#run the following lines together
plot(all_threats,ext=extent(-130,-110,24,50),col=cols) #extent to zoom in
map('world',fill=T,add=T,col='gray')

#for a different extent: Santa Barbara Channel

plot(all_threats,ext=extent(-121,-117,32,35),col=cols, main = "Cumulative Threats") #main adds title

#the zoom() function allows you to draw your extent by clicking twice. run together with plot, but not in markdown html
plot(all_threats, col = cols)


```

Raster data Attributes
```{r}
#frequency histogram od raster cell values

hist(all_threats, main = "Cumulative Threats Frequency")

cellStats(all_threats, mean)
cellStats(all_threats, sd)

```

###Part 3: Raster Calculations

Overlay top 20% of cumulative threats with the top 20% of species richness to find threat hotspots
Steps:
1. Import the data (already done for threats)
2. Resmaple the data to the sample resolution
3. Reclassify the data to find the top 20%
4. Overlay the data and find hotspots

1. Import the data (already done for threats)
```{r}
all_spp <- raster("ca_curr_sp_rich.tif")

#all_spp

plot(all_spp, col = cols)
```

2. Resmaple the data to the sample resolution and extent (Cropping and Resolving)
threats has a greater extent
threats has a finer resolution
```{r}
#crop the threats layer to the extent of the species layer

threats_crop <- crop(all_threats, all_spp) #crops threat layer to same extent of species

#resample species layer to have same resolution as threats. Subsample species layer to arteficially have a higher resolution

spp_res <- resample(all_spp, threats_crop, method = 'ngb', progress = 'text')
# progrerss = 'text' prints out the progress in the console. good for longer running function so you can see progress
# method = 'ngb' specifices that we want to use a nearest neighbor algorithm to resmaple, instead of interpololation


#Check that layers line up by using stack() function
spp_threat_stack <- stack(threats_crop, spp_res)
plot(spp_threat_stack, col = cols)

```

3. Reclassify the data to find the top 20%

first reclassify species
```{r}
hist(spp_res, main = "Species Raster Values")
#this layer has a lot of zeroes --> reassign zeroes which from the plot look like they should be NAs
spp_res <- reclassify(spp_res, rcl = c(-Inf,0,NA)) #overwrites the original spp_res object , -INF = - infinity
hist(spp_res, main = "Species Raster Values, Zeroes removed")

#plot to see what does
#plot(spp_res)

#find the top 20% of the species data, assigning those cells a value of 1, and all other non-NA cells a value of 0. use the quantile() function

spp_cutoff <- quantile(spp_res, 0.8) #value of 80th percentile
spp_maxVal <- cellStats(spp_res, max) #find the maximum

#reclassification matrix
rcl_mat <- c(-Inf, spp_cutoff, 0,
             spp_cutoff, spp_maxVal, 1)

#reclassify species layer
spp_binary <-reclassify(spp_res, rcl = rcl_mat)

#plot to visualize
binary_cols <-c("white", "firebrick") #new color scheme
plot(spp_binary, col = binary_cols, legend = F, main = "Top 20% of Species Richness")
map('world', fill = T, add = T, col = 'gray')


```

now reclassify threats
```{r}

#find the top 20% of the threats data, assigning those cells a value of 1, and all other non-NA cells a value of 0. use the quantile() function

threats_cutoff <- quantile(threats_crop, 0.8) #value of 80th percentile
threats_maxVal <- cellStats(threats_crop, max) #find the maximum

#reclassification matrix
rcl_mat <- c(-Inf, threats_cutoff, 0,
             threats_cutoff, threats_maxVal, 1)

#reclassify species layer
threats_binary <-reclassify(threats_crop, rcl = rcl_mat)

#plot to visualize
plot(threats_binary, col = binary_cols, legend = F, main = "Top 20% of Cumulative Threats")
map('world', fill = T, add = T, col = 'gray')

```


4. Overlay the data and find hotspots
```{r}
#overlay layers to find hotpsots using the overlay() function

hotspots <- overlay(spp_binary, threats_binary, fun = function(x,y){x + y}) #can just add them together because the rasters we are combining are just 1s and 0s

#color breakpoints. We need three colors now (0,1,2)
#brks_hotspots <- seq(0,3,length.out = 4) #not used
hotspot_cols <- c("white", "lightblue", "firebrick")

plot(hotspots, col = hotspot_cols, legend = F, main = "Hotspots"); map('world', fill = T, add = T, col = 'gray80')

data("World")

all_hotspots2 <- reclassify(hotspots, rcl = c(-Inf, 1.5, NA))
hotspot_cols2 <- c("firebrick")

all_hotspot_map <- tm_shape(all_hotpots2) +
  tm_raster("layer", title = " ", palette = hotspot_cols2, labels = "All Threats Hotspots") +
  tm_shape(World) +
  tm_fill("gray80") +
  tm_borders("black") +
  tm_layout(bg.color = "lightblue",
            legend.position = c(0.05, .15),
            legend.text.color = "black",
            legend.text.size = .5, 
            legend.width = 5) +
  tm_compass(north = 0, type = "arrow", position = c("right", "top")) +
  tm_credits("Map: Callie Steffen and Priscilla Hare, June 1, 2019\nSource: Halpern et al 2009\nPublisher: ESM 270", position = c(0.05, .05), size = .4)

all_hotspot_map

tmap_save(all_hotspot_map, "all_hs_map.png", height = 5)




#zoom in SB channel
plot(hotspots, col = hotspot_cols, ext = extent(-121,-117,32,35), legend = F, main = "Hotspots, SB Channel"); map('world', fill = T, add = T, col = 'gray80')

```

